National Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs

### 2020-11-27 Manuscript_Bayesian.Rmd
changes to manuscript.

---
title             : "National Longitudinal Evidence for Growth in Subjective Wellbeing from Spiritual Belief"
shorttitle        : "Longitudinal Wellbeing from Spiritual Beliefs"
author: 
  - name          : "Benjamin Highland"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "520 Power Plant Circle Apartment 105 Winston-Salem, NC 27101 USA"
    email         : "benhighland6420@gmail.com"
  - name          : "Everett L. Worthington"
    affiliation   : "2"
    institution   : "Department of Psychology Virginia Commonwealth University"
  - name          : "Don E Davis"
    affiliation   : "3"
    institution   : "Department of Psychology Georgia State University"
  - name          : "Chris G. Sibley"
    affiliation   : "4"
    institution   : "School of Psychology University of Auckland"
  - name          : "Joseph A. Bulbulia"
    affiliation   : "5"
    institution   : "School of Psychology Victoria University"

@@

keywords          : "Belief, God, Health, Longitudinal, Panel, Religion, Spirit, Spirituality, Well-being"
wordcount         : "4685 words in text body, 1191 words in reference section"
bibliography      : ["b-references.bib"]
floatsintext      : no
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
extra_dependencies: ["rotating", "booktabs", "graphicx","longtable","amsmath","amsfonts","amssymb"]

>>>>

---
title             : "National Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs"
shorttitle        : "Longitudinal Wellbeing from Spiritual Beliefs"
author: 
  - name          : "Benjamin Highland"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "520 Power Plant Circle Apartment 105 Winston-Salem, NC 27101 USA"
    email         : "benhighland6420@gmail.com"
  - name          : "Everett L. Worthington"
    affiliation   : "2"
    institution   : "Department of Psychology Virginia Commonwealth University"
  - name          : "Don E Davis"
    affiliation   : "3"
    institution   : "Department of Psychology Georgia State University"
  - name          : "Chris G. Sibley"
    affiliation   : "4"
    institution   : "School of Psychology University of Auckland"
  - name          : "Joseph A. Bulbulia"
    affiliation   : "5"
    institution   : "School of Psychology Victoria University"

@@

keywords          : "Belief, God, Health, Longitudinal, Panel, Religion, Spirit, Spirituality, Well-being"
wordcount         : "4685 words in text body, 1191 words in reference section"
bibliography      : ["b-references.bib"]
floatsintext      : no
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
extra_dependencies: ["rotating", "booktabs", "graphicx","longtable","amsmath","amsfonts","amssymb"]

### 2020-11-27 Manuscript_Bayesian.tex
changes to manuscript.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={National Longitudinal Evidence for Growth in Subjective Wellbeing from Spiritual Belief},
  pdfauthor={Benjamin Highland1, Everett L. Worthington2, Don E Davis3, Chris G. Sibley4, \& Joseph A. Bulbulia5},
  pdflang={en-EN},
  pdfkeywords={Belief, God, Health, Longitudinal, Panel, Religion, Spirit, Spirituality, Well-being},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{National Longitudinal Evidence for Growth in Subjective Wellbeing from Spiritual Belief}
\author{Benjamin Highland\textsuperscript{1}, Everett L. Worthington\textsuperscript{2}, Don E Davis\textsuperscript{3}, Chris G. Sibley\textsuperscript{4}, \& Joseph A. Bulbulia\textsuperscript{5}}
\date{}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={National Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs},
  pdfauthor={Benjamin Highland1, Everett L. Worthington2, Don E Davis3, Chris G. Sibley4, \& Joseph A. Bulbulia5},
  pdflang={en-EN},
  pdfkeywords={Belief, God, Health, Longitudinal, Panel, Religion, Spirit, Spirituality, Well-being},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{National Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs}
\author{Benjamin Highland\textsuperscript{1}, Everett L. Worthington\textsuperscript{2}, Don E Davis\textsuperscript{3}, Chris G. Sibley\textsuperscript{4}, \& Joseph A. Bulbulia\textsuperscript{5}}
\date{}

### 2020-11-27 Supplement_And_Analysis.Rmd
changes to manuscript.

---
title             : "Supplement: Analysis for National Longitudinal Evidence for Growth in Subjective well-being from Spiritual Belief"
author: "contact: joseph.bulbulia@gmail.com on behalf of Highland et al."
date: "27/11/2020"
authornote: |
  <!-- https://tinyurl.com/ybremelq -->
# keywords          : "keywords"
wordcount         : no
bibliography      : ["b-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
# documentclass     : "apa6"
classoption       : "man"
output            : "html_document" 
editor_options: 
  chunk_output_type: console
---

@@

```{r global_options, echo=FALSE, include = FALSE}
knitr::opts_chunk$set(fig.path="Figs/", 
                      message=FALSE, 
                      warning=FALSE,
                      collapse =TRUE,
                      echo=TRUE, 
                      #results="hide", 
                      fig.width= 10,
                      fig.height=8,
                      tidy = "styler")
```

@@

```{r latex_data_table, include=FALSE}
# you might need to install the dev verion.//remotes::install_github("tysonstanley/furniture")
demographic_table_latex(d_3) # note the furniture package conflicts with the table1 package. 
conflicts()
furniture::table1(d_3, splitby = ~Wave,
                    Age, 
                    NZdep,
                    Education,
                    Employed,
                    Ethnic_Categories,
                    Male,Has_Partner, 
                    Political_Conservativism,
                    Urban, 
                    Beliefs, 
                    LIFESAT,
                    PWI,
                    overall=F, 
                    output = "latex2", 
                    booktabs=TRUE)
```

@@

In the the reported analyses, we included covariates to adjust for potential confounding effects from usual suspect factors associated with our two measures of well-being (Personal Well-Being, Life Satisfaction). Where $Z$ is a third variable, $X$ is the predictor, and $Y$ is the outcome, there are at least four types of confounding that might arise: 
1. $Z$ causes $X$ and  $Z$ causes $Y$: $Z \leftarrow Z \rightarrow Y$
2. $X$ causes $Z$ and  $Z$ causes $Y$: $X \rightarrow Z \rightarrow Y$
3. $X$ causes $Z$ and $Y$ causes $Z$: $X \rightarrow Z; Y \rightarrow Z$
4. $X$ causes $Z$; $Z$ causes $D$; and we condition on D (the descendent of $Z$: $Y$ causes $Z$: $Z$ \rightarrow D; $X \rightarrow Z; Y \rightarrow Z$ (see: [@mcelreath2020statistical], Chapter 6).

@@

Below we write the model in a notation that explicitly states the the default weakly informative priors that the BRMS package produces. And indicated in the model we include group-level indicator for variation in subject-level intercepts $\alpha_j$, a group-level indicator for subject-level time (or "growth")  slope  $\Beta_j$, and we model the co-variance of $\alpha_j$, Below we write the model in a notation that explicitly states the the default weakly informative priors that the BRMS package produces.  And indicated in the model we include group-level indicator for variation in subject-level intercepts $\alpha_j$, a group-level indicator for subject-level time (or "growth")  slope  $\Beta_j$, and we model the co-variance of $\alpha_j$ and  $\Beta_j$ which for computational reasons is estimate as a correlation matrix of (${\bf S}$} and ${\bf R}$).

@@

```{r texreg_bayes_table:pwi,cache=TRUE, include=FALSE}
#Results for publication.
table_model_latex_pwi(mod.1_b)
```

```{r texreg_bayes_table:ls, cache=TRUE, include=FALSE}
table_model_latex_ls(mod.2_b)
```

@@

```{r mixing_sd_pwi}
plot(mod.1_b, pars ="sd_") # requires stronger priors
```

```{r mixing_ls_pwi}
plot(mod.2_b, pars ="sd_") # requires stronger priors
```

>>>>

---
title             : "Supplement: Analysis for National Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs"
author: "contact: joseph.bulbulia@gmail.com on behalf of Highland et al."
date: "27/11/2020"
authornote: |
  <!-- https://tinyurl.com/ybremelq -->
# keywords          : "keywords"
wordcount         : no
bibliography      : ["b-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
# documentclass     : "apa6"
classoption       : "man"
output            : "html_document" 
editor_options: 
  chunk_output_type: console
---

@@

```{r global_options, echo=FALSE, include = FALSE}
knitr::opts_chunk$set(fig.path="figs/", 
                      message=FALSE, 
                      warning=FALSE,
                      collapse =TRUE,
                      echo=TRUE, 
                      #results="hide", 
                      fig.width= 10,
                      fig.height=8,
                      tidy = "styler")
```

@@

<!-- ```{r latex_data_table, include=FALSE} -->
<!-- # you might need to install the dev verion.//remotes::install_github("tysonstanley/furniture") -->
<!-- #demographic_table_latex(d_3) # note the furniture package conflicts with the table1 package.  -->
<!-- ``` -->

@@

In the reported analyses, we included covariates to adjust for potential confounding effects from usual suspect factors associated with our two measures of well-being (Personal Well-Being, Life Satisfaction). Where $Z$ is a third variable, $X$ is the predictor, and $Y$ is the outcome, there are at least four types of confounding that might arise: 
1. $Z$ causes $X$ and  $Z$ causes $Y$: $Z \leftarrow Z \rightarrow Y$
2. $X$ causes $Z$ and  $Z$ causes $Y$: $X \rightarrow Z \rightarrow Y$
3. $X$ causes $Z$ and $Y$ causes $Z$: $X \rightarrow Z; Y \rightarrow Z$
4. $X$ causes $Z$; $Z$ causes $D$; and we condition on D (the descendent of $Z$: $Y$ causes $Z$: $Z$ \rightarrow D; $X \rightarrow Z; Y \rightarrow Z$ (see: [@mcelreath2020statistical], Chapter 6).

@@

Below we write the model in a notation that explicitly states the the default weakly informative priors that the BRMS package produces. And indicated in the model we include group-level indicator for variation in subject-level intercepts $\alpha_j$, a group-level indicator for subject-level time (or "growth")  slope  $\beta_j$, and we model the co-variance of $\alpha_j$, Below we write the model in a notation that explicitly states the the default weakly informative priors that the BRMS package produces.  And indicated in the model we include group-level indicator for variation in subject-level intercepts $\alpha_j$, a group-level indicator for subject-level time (or "growth")  slope  $\beta_j$, and we model the co-variance of $\alpha_j$ and  $\beta_j$ which for computational reasons is estimate as a correlation matrix of (${\bf S}$} and ${\bf R}$).

@@

<!-- ```{r texreg_bayes_table:pwi,cache=TRUE, include=FALSE} -->
<!-- #Results for publication. -->
<!-- table_model_latex_pwi(mod.1_b) -->
<!-- ``` -->

<!-- ```{r texreg_bayes_table:ls, cache=TRUE, include=FALSE} -->
<!-- table_model_latex_ls(mod.2_b) -->
<!-- ``` -->

@@

```{r mixing_sd_pwi, cache=TRUE}
plot(mod.1_b, pars ="sd_") # requires stronger priors
```

```{r mixing_ls_pwi, cache=TRUE}
plot(mod.2_b, pars ="sd_") # requires stronger priors
```
